Appendix B of Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages

### 2021-04-26 manuscript/bef_stability/appendixB.Rmd
Minor edits
Updates

---
title: Appendix B of Species richness and network structure jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   -  \newcommand{\beginsupplement}{\setcounter{table}{2}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

The fish community sampling took place in a variety of stream size, shape and
depth. This diversity constraints the type of protocol that can take place.  For
the small streams (depth < 0.7m across the entire width), the complete protocol
is used. It consists in placing anodes across the width of the
stream. The fishermen made from one to three row
of fishing. We kept only the first row for analysis to homogenize the sampling
effort. For deeper and wider streams, the sampling was partial over the bank and
the channel and performed by boat and by foot (along the bank). We kept only
sampling done over the bank because channel sampling was not always done. Two
types of partial protocols takes place: partial over bank and partial by points.
The former is continuous while the latter is discrete. Furthermore, in the "partial
by point" protocol, the channel can be sampled as well. We kept only the "partial
by point" sampling in which all the sampling points were done over the bank to
homogenize the sampling. Then, we merged partial over bank and partial by 
point protocols. Some sampled stations had mixed characteristics in which mix of
partial and complete protocols were applied depending of the depth of the stream 
at the time of sampling. In this case, we kept only the protocol that was the
most used.

Heterogeneity in the sampling period may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in sampling time, only one sampling per year was retained if there
was two in a given year, the one with the most species and then individual
collected. Finally, most of the fish sampling occurred in September (Median:
mid-August) and the median standard deviation of the sampling month was inferior
to one (0.7) within stations. Additionally, we removed sampling events, in a
given station, in which the sampled length along the stream was outside the median
length sampled more or less 30%. Finally, we standardized the biomass record
and species richness by the sampled surface to get rid of sampling effort
effects.

@@

During the electric fishing, and before that fish body size are measured, the
fishes are regrouped by batches. Each batch contains one species but there can
be several batches by species. Body size assessment are different according to
the batches. There are four types of batches: N, I, S/L and G. The batches N and
I contains the biggest individuals which are normally less numerous. Then
the body size of all the individuals in the N and I batches is recorded. The batches S/L
contains fishes of intermediate sizes. Inside each S/L batches, the operator
tries to make batches as homogeneous as possible in terms of body size. Then, the body
size of a subsample of at least 20 individuals is recorded. The G batches
contain individuals of small body size, with care that body size is as
homogeneous as possible in a given batches. Then body size of the shortest and the
longest individuals are recorded.

@@

With the information contained in the batch record, we inferred the body size of
the fishes in the community. No inference were needed for the N and I batches
because the body size of all individuals was recorded. The body size ($l$) of the
fishes in the S/L and G was inferred under the assumption that its distribution
followed a normal law ($l ~ N(\mu , \sigma)$).
For a S/L batch, the mean $\mu$ and the standard deviation $\sigma$ were simply
the average and the standard deviation of the body size of the subsample
recorded. For a G batch, the average was approximated by 
$\hat{\mu} = (l_{min} + l_{max}) 2^{-1}$ and the standard deviation by
$\hat{\sigma} = (l_{min} + l_{max}) 4^{-1}$. Together with the estimations of
the mean and the standard deviation, the number of individuals in a batch and
assumption of a normal distribution, we simulated the body size distribution in
S/L and G batches. To avoid simulation of unrealistic body size, we used a
truncated normal distribution bounded at [0.05; .95] for the density of
probability of the S/L batches, and bounded at [$l_{min}$; $l_{max}$] for the G
batches. Body size distribution were generated using the R function
`truncdist::rtrunc`.

We removed batches that did not satisfied the following conditions. First, we
removed batches in which the total number of individuals was not recorded. In lot
"G", we removed batches where $l_{min} \geq l_{max}$ or if at least one of the
measurement was missing. Additionally, G batches were removed when number of
individuals fell below five, as G batches normally contained more individuals.
In the S/L lot, we removed batches in which the number of individuals in the
recorded individuals fell below 10. In sum, those sanity check allowed to
exclude potentially problematic batches. Finally, in the dataset used in the
analysis, `r na_length` individuals had missing length over 
`r nrow_fish_length` individuals.

@@

We removed rare fish species, species from lake, crayfish species, and migratory
species. In total, we removed `r nrow(removed_ind)` over 
`r nrow(length_sem_analysis)` individuals and `r length(removed_sp)` species.

During fish sampling, some species are identified based on morphotype but they
belong to the same species based on fecundation criteria.  We merged those
morphotype species into one species. There was 
`r length(unique(names(sp_to_replace())))` species to merge which belonged to 
`r length(unique(sp_to_replace()))` species.

@@

```{r}
trophic_building_caption <- paste0(
  "Trophic species"
) 
```

```{r p-net-inf, fig.cap=trophic_building_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "network_inference.pdf"))
```

@@

The metaweb described the potential trophic interactions between the
($`r ncol(metaweb_analysis$metaweb) - length(metaweb_analysis$resource)`$) trophic species.
We had seven resource nodes which were present in the diet of the trophic
species namely detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos.

The metaweb building was based on trait matching, so it necessitates to have
information about their diet. `r length(removed_sp_metaweb)` species were
removed because information about their diet was missing. It represented  
`r nrow(removed_ind_metaweb)` individuals over `r nrow(length_sem_analysis)` 
individuals.

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies [@brose_predator_2019; @poisot_structure_2016;
@schneider_body_2012].  For a given trophic species, we first determined if its
size class ends up in a life stage containing piscivorous diet, i.e. if there was
an overlap between them. We were more liberal for piscivory determination to
avoid 
missing potential fish-fish trophic interactions which were the main interactions
in our community. If the trophic species was piscivorous, we determined its prey
size range from 3% to 45% of the median of its size range. A trophic interaction
was set for each trophic species whose median of its size class fell into the
prey range of the predator trophic species. To sum up, the metaweb was obtained
with allometric feeding interactions (1) between trophic species, (2) between
trophic species and resource nodes and (3) between resource nodes.

@@

The local trophic network (i.e. one network for each fish community
sampling) was obtained by determining the trophic species of each fish
individual in the community and interactions were set according to the metaweb.
All the resource nodes were assumed to be present in each local network.
Finally, the structure of the local networks varied according the presence of
the trophic species. The inferred interactions represent potential interactions
(i.e. not measured interactions) and are binary (i.e. presence or absence of
feeding interactions).

@@

The procedure to obtain altitude, slope, width depth of the streams are fully
explained in main text. Because those measurements were collected during the
fishing protocol, they were easier to manage in the analysis. Then, we detailed
in the following sections how water temperature, flow and BOD were obtained.

@@

Water temperature, flow and BOD (Biological Oxygen Demand) were obtained from the
[naiades database](http://www.naiades.eaufrance.fr/).
Water temperature, flow and BOD were respectively collected at a hourly, daily
and monthly frequency. Water temperature and flow are recorded by automated
sensors while BOD is collected manually because it necessitates analysis in a
laboratory. BOD data collection is coordinated by the Agence de l'eau which
realized the analysis itself or order it to a company.
It is crucial to make clear that those variables are collected at different
places than the fish samplings.

```{r}
pressure_cap <- paste0(
  "Map of environmental variables obtained for the 99 sites used in the analysis."
)
```

```{r, fig.height = 20/2.54, fig.cap = pressure_cap}
pca_pressure <- habitat_press
myload(fourier, dir = mypath("data-raw", "polluants"))
myload(flow_temp_metrics, station_analysis, the_8_hydrologic_basin, dir = mypath("data"))

@@

The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each
recording station, we applied a moving window averaging to get rid of seasonal
effects. For temperature, the moving was performed on daily average temperature.
Then we took the annual average of each year and each variable (i.e. one value
by year and by station).

Several characteristics of the streams make the interpolation more complicated
than along terrestrial surface. Interpolations in terrestrial areas are roughly
based on euclidean distance. Euclidean distance is more complicated to consider
in streams because it had the follow the path of the streams. Furthermore, two
streams can be close by but not connected and the value of their variables could
be not related or a only a few. For those reason, we used stream spatial network
methods to interpolate water temperature, flow and BOD.

@@

The procedure was done by hydrographic basin because of convergence problem
when the interpolation was done at the France scale. We had to make a trade-off
between the number of data (to keep enough high to perform the interpolation)
and the level of dataset splitting. We then merge some hydrographic basin to end
up with four region of interpolation: West, East, South and North.

Then we modeled each variable as dependant only on the covariation between sites
and taking in account stream network. To model the autocovariance between sites,
we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their prediction performance. The interpolations
were done with the `SSN` R package 
[@isaak_applications_2014; @hoef_ssn:_2014].

@@

```{r eval = TRUE}
myload(cv_press_interp_mv_avg,
  dir = mypath("data-raw", "polluants"))
myload(cv_temp_press_interp_mv_avg,
  dir = mypath("data-raw", "naiades_temperatures"))
myload(cv_flow_press_interp_mv_avg,
  dir = mypath("data-raw", "flow"))
dbo_cv <- cv_press_interp_mv_avg %>%
  filter(parameter == "dbo5")

@@

cv_dbo_temp_flow %>%
  mutate_if(is.double, ~signif(., 2)) %>%
  kable(
    align = "c",
    format = "latex",
    booktabs = T,
    caption = paste0("Results of the cross-validation for the interpolation of
      BOD, temperature and flow. Average over 4 interpolations are shown.") 
      ) %>%
    collapse_rows(columns = c(1, 2), valign = "top") %>%
    kable_styling(latex_options =c("striped", "scale_down"))
```

```{r}
cv_bias_caption <- paste0(
  "Bias of the interpolation from Leave-One-Out Cross Validation (LOOCV). "
)
```

The performance of the interpolation was evaluated with leave one out
cross-validation (already implemented in the SSN package), i.e. the value of the
interpolated variable of each station was predicted after removing it from the
dataset.

When predicted values were below the minimum value or above the maximum value
recorded in the dataset for flow, temperature and BOD, we replaced the
predicted value by respectively the minimum and the maximum value found in the
dataset. We ran the interpolation four times and we kept the predicted value
which had the least standard deviation.

@@

The CV of biomass over time can cover multiple dynamics, including increases,
stable, decreases and a mix of them. Theoretically, however, the CV computed to
quantify stability is often represented with the assumption that the overall
dynamic of the biomass varies around an equilibrium. In order to approach this
condition, we kept only the stations that have met this criteria. To qualify the
biomass trajectories over time, we used the symbolic aggregated approximation
method (SAX) which is commonly used in time series analysis and implemented in
the `jmotif` R package. This method classifies time series rather than
quantifies trends and so is generic enough to encompass the diversity of time
series shapes. The classification necessitates three steps. First, the time
series are normalized (i.e. substituting the mean and divide by the standard
deviation). The second step is to reduce the dimensionality of the time series
by Piecewise Aggregation Approximation (PAA) method. Each time series is divided
in $N$ frames of equal size that allows to the reduce the dimensionality of the
time series. For each frame, we simply kept the mean value of the time series.
The last step is to qualify each frame into string. Because values of
normalized time series typically follows a normal distribution ($N(0,1)$), a
breakpoints can be computed by dividing the area under the normal distribution
in equal parts. The algorithm attributes a alphabetic letter to each part from
the lowest to the highest ($a$, $b$, $c$, ...). We set three breakpoints, so
$a$, $b$ and $c$. Finally, the trajectory of each time series is classified as
a combination of $a$, $b$, $c$ letters. $abc$ and $cba$ time series represent
respectively monotonic increasing and decreasing trajectories. For data analysis,
we kept only the $bbb$ which represent time series that showed no temporal
tendencies, with respect to our methodology.

@@

```{r contrib-pca}
#https://stats.stackexchange.com/a/143949
caption_contrib_pca <- paste0(
  "Loadings of variable to rotated components of the PCA.
  Loadings are the coefficients which determine the components of the PCA as
  linear combinations of the variables.
  "
)

@@

```{r, eval=FALSE}
corr_pca_caption <- paste0(
  "Spearman correlation between environmental variables used in the PCA.
  The variables have been scaled before to compute the correlation."
)

@@

pca_caption <- paste0(
  "Rotated axis of PCA on environmental variables. The five components
  explained,", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  "(A) Axis 1 was positively related to the stream size, axis 2 was related to
  the stream slope and altitude and negatively related to average water temperature and
  BOD. (B) Eigenvalues distribution, we kept the first two axis for the data analysis. ",
  "Please note that we took the inverse of the second axis in the data analysis,
  to make the axis more easily interpretable, i.e. positively related de average
  water temperature."

  )
```

@@

pca_plot <- plot_grid(p1, eigen_rotated_plot, nrow = 1)
save_plot(
  filename = mypath("manuscript", "bef_stability", "figs", "pca_plot.pdf"),
  pca_plot,
  nrow = 1, ncol = 2,
  base_height = 6/2.54, base_asp = 1.6, base_width = NULL 
  )
pca_plot
```

@@

```{r}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 2)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 2), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c("Log base 10", "Log base 10 and inversion", "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  interpretation = c("linked to stream size",
    "linked to average temperature and BOD",
    "Total species richness over the temporal series",
    "Median of annual link density", "Median of annual average trophic level", "", "",
    "Median of annual total biomass")) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "Variable description and transformation"
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
sem_variable_table
```

@@

```{r}
hist_var_caption <- paste0(
  "Distribution of the variables used in the Structural Equation Modeling. The
  biomass stability, the CV of species (CVsp), PCA1, PCA2, species richness,
  synchrony and total biomass were logged."
)
```

>>>>

---
title: Appendix B of Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages 
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \newcommand{\beginsupplement}{\setcounter{table}{2}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
nocite: |
    @post_ecosystem_2000| @doi_resource_2009| @mchugh_dual_2010|
    @ogorman_unexpected_2017| @gauzens_biodiversity_2020| @dunne_network_2006|
    @mchugh_dual_2010| @doi_resource_2009| @allan_stream_2009|
    @petchey_predicting_2010| @winemiller_must_1989
csl: ecology-letters.csl
---

@@

The monitoring of the fish communities occurs in a variety of stream size, shape
and depth. The stream characteristics constrain the type of protocol that can be
applied. Small streams (depth < 0.7m across the entire width) are sampled with
the complete protocol. The protocol consists of placing anodes across the entire
width of the stream. The fishermen made from one to three rows of electric
fishing. We selected only the data from the first row to homogenize the sampling
effort. In deeper and wider streams (depth >= 0.7m), the sampling was partial
over the bank and the channel and performed by boat and by foot (along the
bank). We selected only the sampling done over the bank because the channel was
not systematically sampled. Two types of partial protocols takes place: partial
over bank and partial by points, which are respectively continuous and discrete.
Furthermore, in the "partial by point" protocol, the channel was sometimes
sampled as well. We selected only the "partial by point" sampling in which all
the sampling points were done over the bank to homogenize the sampling. Then, we
merged partial over bank and partial by point protocols. Some sampled stations
were located in stream with intermediate characteristics where both protocols
were alternatively used depending on the state of the stream at the time of
sampling. For these sites, we only kept the samples that had been carried out
with the most used protocol.

Heterogeneity in the sampling date may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in the sampling date, only one sampling per year was retained if
there was more in a given year, the one with that had the most species and then
the most individuals. Finally, most of the fish sampling occurred in September
(Median: mid-August) and the median standard deviation of the sampling month was
inferior to one (0.7) within stations, i.e. 3 weeks. To further reduce sampling
heterogeneity, we only kept the sampling in which the length of stream sampled
along the stream was inside the median length sampled more or less 30%.

@@

Once the fishes have been caught, they are divided into different batches (Fig.
\@ref(fig:p-body-size)). Each batch contains only one species but there may be
several batches for one species. The body size of the fishes is measured
differently for each type of batch.  There are four types of batches: N, I, S/L
and G. The batches N and I contain the biggest individuals. In the N and I
batches, the body size of all the individuals is recorded. The batches S/L
contain fishes of intermediate sizes.  Inside each S/L batch, the operator tries
to make batches as homogeneous as possible in terms of body size. Then, the body
size of a subsample of at least 20 individuals is recorded. The G batches
contain individuals of small body size, with care that body size is as
homogeneous as possible in a given batches.  Then body size of the shortest and
the longest individuals are recorded.

@@

We inferred the body size of all the unmeasured individuals using the batch
records. There was no need for inference for the N and I batches since the body
size of all individuals was recorded (Fig. \@ref(fig:p-body-size)). The body
size ($l$) of the fishes in the S/L and G was inferred under the assumption that
body size distribution follows a normal law ($l ~ N(\mu, \sigma)$).  For an S/L
batch, the mean $\mu$ and the standard deviation $\sigma$ were simply the
average and the standard deviation of the body size of the subsample recorded.
For a G batch, the average was approximated by $\hat{\mu} = (l_{min} + l_{max})
2^{-1}$ and the standard deviation by $\hat{\sigma} = (l_{min} + l_{max})
4^{-1}$. With the estimations of the mean and the standard deviation, the number
of individuals in a batch and the assumption of normal distribution, we
simulated the body size distribution in S/L and G batches. To avoid simulation
of unrealistic body size, we used a truncated normal distribution bounded at
[0.05; .95] for the density of probability of the S/L batches, and bounded at
[$l_{min}$; $l_{max}$] for the G batches. Body size distribution were generated
using the R function `truncdist::rtrunc`.

Batches that did not meet certain conditions were excluded. First, we also
excluded batches for which the total number of individuals was not recorded. We
also excluded G batches for which $l_{min} \geq l_{max}$ or if
at least one of the measurement was missing. Additionally, a given G batch was
excluded when the number of individuals fell below five, as G batches normally
contained more individuals. We finally excluded S/L batches for which the
number of measured individuals was inferior to 10. Finally, in the dataset used
in the analysis, `r na_length` individuals had missing length over 
`r nrow_fish_length` individuals.

@@

The rare fish species, species from lake, crayfish species, and migratory
species were excluded. In total, we removed `r nrow(removed_ind)` over 
`r nrow(length_sem_analysis)` individuals and `r length(removed_sp)` species.

Some species were identified based on their morphotype, but they
belong to the same species based on fecundation criteria. We merged those
morphotype species into one species. There was
`r length(unique(names(sp_to_replace())))` species to merge which belonged to 
`r length(unique(sp_to_replace()))` species.

@@

```{r, cache = FALSE}

trophic_building_caption <- paste0( "Building of the metaweb. (1) Each
  species are divided in nine body size classes. We gathered information about
  the food diet of each fish species and how food diet varies according to
  ontogeny. (2) The life stage of a trophic species was determined with the
  median of the body size class. A trophic link was set between the trophic
  species and the resources associated to this life stage. (3) For a piscivorous
  trophic species, the trophic links between fishes were defined according to
  the predators/prey body size ratio. The predation window of the trophic
  species ranged from 3% to 45% of the median of the body size range. A trophic
  link between the predator trophic species and every trophic species whose
  midpoint of bodysize range was included in the predation window.")

```

```{r p-net-inf, cache = FALSE, fig.cap=trophic_building_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "network_inference.pdf"))
```

@@

```{r, cache = FALSE}
metaweb_caption <- paste0(
  "The metaweb describes the trophic interactions between all the 412 trophic species, each is
  represented by a node and each color represents a species (S = 45). The seven
  resource nodes (7) are drawn in grey. The species names corresponding to the
  three digit codes are found in Table S1."
)
```

```{r metaweb, cache = FALSE, fig.cap=metaweb_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "metaweb.png"))
```

The metaweb describes the potential trophic interactions between the
($`r ncol(metaweb_analysis$metaweb) - length(metaweb_analysis$resource)`$) trophic species.
The metaweb also included seven resource nodes which were present in the diet of the trophic
species namely detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos (Fig. \@ref(fig:metaweb)).

Trophic interactions between the nodes of the metaweb were determined according to
the diet and the body size of the trophic species. `r length(removed_sp_metaweb)` 
species were removed because information about their diet was missing. It
represented  `r nrow(removed_ind_metaweb)` individuals over
`r nrow(length_sem_analysis)` individuals.

The life stage to which a trophic species belongs was determined by the median of
the body size range of the trophic species, each life stage being delimited by a body
size range. In turn, this trophic species has trophic interactions with the
resource nodes of its life stage. The trophic interactions between a
predator and a prey were determined by predator-prey body size ratio
[@brose_predator_2019; @poisot_structure_2016; @schneider_body_2012]. For a
given trophic species, we first determined if its body size range overlap
with a piscivorous life stage (Fig. \@ref(fig:p-net-inf)). We were then more
liberal for piscivory determination than for fish-resource interactions, to
avoid missing potential interactions which were the main interactions in our
community. If the trophic species was piscivorous, we determined its prey size
range from 3% to 45% of the median of its size range. A trophic interaction was
set for each trophic species whose median of its size class fell into the prey
range of the predator trophic species. To sum up, the metaweb contains feeding
interactions (1) between trophic species, (2) between trophic species and
resource nodes and (3) between resource nodes.

@@

The local trophic networks (i.e. one network by
sampling) were obtained by determining the trophic species of each fish
individual in the community. According to the trophic species present in a given
community, we set the trophic interactions between them according to the metaweb. 
All the resource nodes were assumed to be present in each local network.
Finally, the structure of the local networks varied according the presence of
the trophic species.

@@

The procedure to obtain altitude, slope, width and depth of the streams are fully
explained in main text. Because those measurements were collected during the
fishing protocol, they were easier to manage in the analysis. Then, we detailed
in the following sections how water temperature, water flow and Biological
Oxygen Demand (BOD) variables were obtained. The goal was to obtain an annual
mean for each variable, to match the yearly resolution of the fish monitoring.
Furthermore, those variables are measured at different locations than the fish
monitoring.

@@

Water temperature, water flow and BOD variables were obtained from the [naiades
database](http://www.naiades.eaufrance.fr/) and were respectively collected at
an
hourly, daily and monthly frequency. Water temperature and flow are recorded by
automated sensors while BOD is collected manually because its measurement
requires laboratory analysis. BOD data collection is coordinated by the Agence
de l'eau which realized the analysis itself or order it to a company.

```{r, cache = FALSE}
pressure_cap <- paste0(
  "Map of the environmental variables obtained for the 99 sites used in the
  analysis. Slope is expressed in degree, altitude in meter, distance from
  source in kilometer, Strahler order has no units, average stream width and
  depth are expressed in meter, Biological Oxygen Demand (BOD) is expressed in
  mg of dissolved oxygen, water flow is expressed in cube meter per second and
  water temperature is expressed in Celsius degree.")
```

```{r, cache = FALSE, fig.height = 20/2.54, fig.cap = pressure_cap}
pca_pressure <- habitat_press
myload(fourier, dir = mypath("data-raw", "polluants"))
myload(flow_temp_metrics, station_analysis, the_8_hydrologic_basin, dir = mypath("data"))

@@

To get rid of seasonal effects of temperature, we applied a moving window
averaging to each time series. The moving window was applied on daily average
temperature. Then we took the annual mean of each year and each variable
(i.e. one value by year and by station).

Several characteristics of the streams make the interpolation more complicated
than along terrestrial surface. Interpolations in terrestrial areas are roughly
based on euclidean distance while euclidean distance is complicated to consider
in streams because it had to follow the path of the streams. Furthermore, two
streams can be close by but not connected and thus the value of their variables might 
not be related or only a few. For those reason, we used stream spatial network
methods to interpolate water temperature, water flow and BOD.

@@

The interpolations were carried out by hydrographic basin due to convergence
issues when the interpolation was carried out at the scale of the whole
France. We had to find a trade-off between the number of data (to keep enough
high to perform the interpolation) and the level of dataset splitting. We then
merged some hydrographic basin to end up with four regions of interpolation:
West, East, South and North.

Then we modelled each variable as dependant only on the covariation between
sites and taking in account stream network. To model the autocovariance between
sites, we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their predictive performance. The interpolations
were done with the `SSN` R package [@isaak_applications_2014; @hoef_ssn:_2014].

@@

```{r, cache = FALSE}
cv_bias_caption <- paste0(
  "Bias of the interpolation from Leave-One-Out Cross Validation (LOOCV). "
)
```

```{r, eval = TRUE}
myload(cv_press_interp_mv_avg,
  dir = mypath("data-raw", "polluants"))
myload(cv_temp_press_interp_mv_avg,
  dir = mypath("data-raw", "naiades_temperatures"))
myload(cv_flow_press_interp_mv_avg,
  dir = mypath("data-raw", "flow"))
dbo_cv <- cv_press_interp_mv_avg %>%
  filter(parameter == "dbo5")

@@

```
```{r, cache = FALSE}

cv_dbo_temp_flow %>%
  mutate_if(is.double, ~signif(., 2)) %>%
  kable(
    align = "c",
    format = "latex",
    booktabs = T,
    caption = paste0("Results of the cross-validation for the interpolation of
      BOD, temperature and flow. Average over 4 interpolations are shown. The
      cross-validation were performed with the SSN R package. bias: average
      predicted error, std.biais: bias standardized, RMSPE: Root Mean-Squared
      Prediction Error, std.MSPE: standardized Mean-Squared Prediction Error.") ) %>%
    collapse_rows(columns = c(1, 2), valign = "top") %>%
    kable_styling(latex_options =c("striped", "scale_down"))
```

The performance of the interpolations was evaluated with leave one out
cross-validation (already implemented in the SSN package), i.e. the value of the
interpolated variable of each station was predicted after removing it from the
dataset.

When predicted values were below the minimum value or above the maximum value
recorded in the dataset for flow, temperature and BOD, we replaced the
predicted value by respectively the minimum and the maximum value found in the
dataset. We ran the interpolation four times, and we kept the predicted value
which had the least standard deviation.

@@

The CV of biomass over time can cover multiple temporal dynamics, including
increases, stable, decreases and a mix of them. Theoretically, however, the CV
computed to quantify stability is often represented with the assumption that the
overall dynamic of the biomass varies around an equilibrium
[@kefi_advancing_2019]. In order to approach the condition of equilibrium, we kept only the
stations that have met this criteria. To characterise the biomass temporal
dynamics, we used the symbolic aggregated approximation method (SAX), which is
commonly used in time series analysis and implemented in the `jmotif` R package
[@senin_grammarviz_2018].  This method classifies time series rather than
quantifies trends and so is generic enough to encompass a diversity of temporal
trajectories. The classification necessitates three steps. First, the time
series are normalized (i.e. substituting the mean and divide by the standard
deviation). The second step is to reduce the dimensionality of the time series
by a Piecewise Aggregation Approximation (PAA) method. Each time series is
divided in $N$ frames of equal size that allows to reduce the dimensionality
of the time series. For each frame, we simply kept the mean value of the time
series. The last step is to qualify each frame into string. Because values of
normalized time series typically follows a normal distribution ($N(0,1)$), a
breakpoint can be computed by dividing the area under the normal distribution
in equal parts. The algorithm attributes an alphabetic letter to each part from
the lowest to the highest ($a$, $b$, $c$, ...). We set three breakpoints, so
$a$, $b$ and $c$.  Finally, the trajectory of each time series is classified as
a combination of $a$, $b$, $c$ letters. $abc$ and $cba$ time series represent
respectively monotonic increasing and decreasing trajectories. For data
analysis, we kept only the $bbb$ which represent time series that showed no
temporal tendencies, with respect to our methodology.

@@

```{r contrib-pca, cache = FALSE}
#https://stats.stackexchange.com/a/143949
caption_contrib_pca <- paste0(
  "Loadings of variable to rotated components of the PCA.
  Loadings are the coefficients which determine the components of the PCA as
  linear combinations of the variables.
  "
)

@@

```{r, eval=FALSE, cache = FALSE}
corr_pca_caption <- paste0(
  "Spearman correlation between environmental variables used in the PCA.
  The variables have been scaled before to compute the correlation."
)

@@

pca_caption <- paste0(
  "Eigenvalues distribution, we kept the first two axis for the data analysis." 
  )
```

@@

pca_plot <- plot_grid(eigen_rotated_plot, nrow = 1)
save_plot(
  filename = mypath("manuscript", "bef_stability", "figs", "pca_plot.pdf"),
  pca_plot,
  nrow = 1, ncol = 2,
  base_height = 6/2.54, base_asp = 1.6, base_width = NULL 
  )
pca_plot
```

@@

```{r, cache = FALSE}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 2)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 2), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c("Log base 10", "Log base 10 and inversion", "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  description = c("linked to stream size",
    "linked to average temperature and BOD",
    "Total species richness over the temporal series",
    "Median of annual link density", "Median of annual average trophic level",
    "Temporal synchrony in the dynamics of population biomass", "Average
    variability of population biomass", "Median of annual total biomass")) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "The description of the variables used in the structural equation models."
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
sem_variable_table
```

\begin{table}
\caption{\label{tab:hyp-sem} Presentation of the links 
of the Structural Equation Models associated with the references supporting the links.}
\begin{center}
\begin{tabular}{ m{7cm}| m{1cm}| m{6cm}}
    \hline
    Link & sign & reference\\
    \hline
    PCA1 Avg stream size $\to$ Species richness & (+) & Post et al. (2000); Doi et al. (2009); McHugh et al. (2010) \\
    \hline
    PCA2 Avg temperature $\to$ Species richness & (+/-) & O’Gorman et al. (2017); Gauzens et al. (2020) \\
    \hline
    PCA1 Avg stream size $\to$ Connectance & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ Connectance & (+/-) & Petchey et al. (2010) \\
    \hline
    Species richness $\to$ Connectance & (-) & Winemiller (1989); Dunne (2006) \\
    \hline
    PCA1 Avg stream size $\to$ Avg trophic level & (+) & Post et al. (2000); Doi et al. (2009); McHugh et al. (2010) \\
    \hline
    PCA2 Avg temperature $\to$ Avg trophic level & (+/-) & O’Gorman et al. (2017)\\
    \hline
    Species richness $\to$ Avg trophic level & (+) & Schneider et al. (2016)\\
    \hline
    Species richness $\to$ Synchrony & (-) & Thébault and Loreau (2005); Loreau and de Mazancourt (2008); Thibaut and Connolly (2013) \\
    \hline
    Connectance $\to$ Synchrony & (+) & Teng and McCann (2004)\\
    \hline
    Avg trophic level $\to$ Synchrony & (+) & Raimondo et al. (2004)\\
    \hline
    PCA1 Avg stream size $\to$ Synchrony & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ Synchrony & (+) & Loreau and de Mazancourt (2008)\\
    \hline
    Species richness $\to$ CVsp & (+) & Tilman (1996) \\
    \hline
    Connectance $\to$ CVsp & (+) & Thébault and Loreau (2005)\\
    \hline
    Avg trophic level $\to$ CVsp & (-) & McCann and Rooney (2009); Shanafelt and Loreau (2018)\\
    \hline
    PCA1 Avg stream size $\to$ CVsp & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ CVsp & (+) & O’Gorman et al. (2017); Tabi et al. (2019)\\
    \hline
    CVsp $\to$ Biomass stability & (-) & Thibaut and Connolly (2013)\\
    \hline
    Synchrony $\to$ Biomass stability & (-) & Thibaut and Connolly (2013)\\
    \hline
    Species richness $\to$ Total biomass & (+) & Schneider et al. (2016)\\
    \hline
    Connectance $\to$ Total biomass & NA & Wang and Brose (2018)\\
    \hline
    Avg trophic level $\to$ Total biomass & (+) & Wang and Brose (2018)\\
    \hline
    PCA1 Avg stream size $\to$ Total biomass & (+) & Post et al. (2000); Allan and Ibanez Castillo (2009); Doi et al. (2009)\\
    \hline
    PCA2 Avg temperature $\to$ Total biomass & (+/-) & O’Gorman et al. (2017)\\
    \hline
    \end{tabular}
\end{center}
\end{table}

@@

```{r, cache = FALSE}
hist_var_caption <- paste0(
  "Distribution of the variables included Structural Equation Models The
  biomass stability, the CV of species (CVsp), PCA1, PCA2, species richness,
  synchrony and total biomass were logged."
)
```

```{r hist-sem-data, cache = FALSE, fig.cap = hist_var_caption}
sem_data %>%
  select(log_bm_std, log_rich_tot_std, log_sync, log_stab_std, log_cv_sp, log_RC1, log_RC2, t_lvl, ct) %>%
  gather(variable, value) %>%
  mutate(variable = str_replace_all(variable, get_sem_var_name_replacement())) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free")
```

```{r, cache = FALSE}
colin_sem_cap <- paste0(
  "Variance Inflation Factor for each individual linear model of the structural
  equation models."
) 
```
